Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-12-30 manuscript.Rmd
Update discussion

///

///

@@

This methodology should be of great interest to interdisciplinary scientists investigating
social mobilization. It has potential applications for gender studies and
the #MeToo movement and for the relationship between social protest and repression
during the Black Lives Matter protests. This methodological
approach can help us better understand the fundamental human social mechanisms underlying
social media and its influence on human behavior in real life. Given that many
governments enact repressive measures when they determine that collective action
by citizens is imminent, this methodology should help to improve the predictive
power of using social media data in estimating repressive behavior. If social
media messages are increasing in cohesion, this may represent a threat to civil
and political liberties of citizens engaging in collective action.

## Conclusion
Considering the pervasiveness of social media, it is unsurprising that social
activists and protesters have utilized these platforms to support real-world
mobilization. Here, we quantitatively explored the connection between online
social cohesion and real-world events. In addition to introducing political
scientists to a family of powerful nonlinear dynamics analyses, our results have
serious implications for monitoring global movements. While state-run media in
regions of authoritarian rule are often biased and serve as propaganda for the
leader and central government, the self-organized nature of Twitter provides a
new avenue to gain insight into truths about uprisings and social mobilization.
By conceptualizing Twitter and real-world events as two elements of an
inextricably linked dynamical system, the levels of social cohesion within
social media platforms could be monitored for fluctuations indicating shifts in
power and peace. Social media mobilization may thus help alert the world to
citizens' actions and reactions that are suppressed by government regimes,
providing insight into events that may not otherwise be broadcast and allowing
for faster intervention.

>>>>

For positive events, we see a similar early peak in RR, but the later
above-chance values did not match up temporally with that of the all-events time
series (FIX Figure 10). This suggests that the early stretch of above-chance RR
may have been captured in the RR of the all-events analysis, whereas the shorter
stretches were overshadowed by the coupling with negative events. The early
period of above-chance DET in the positive-event analysis aligned with the early
period of above-chance RR, whereas the later stretch of significant DET values
aligns temporally with that of the all-events analysis. It is possible that the
later stretch reflected the decrease in the number of positive events happening
(FIX see Fig. 4), while the relative sparsity of positive events led to an
increase in coupling, with denizens rallying around dwindling bits of hope as
time went on.

For negative events, RR rarely reached above-chance values, but around the later
windows, there were above-chance peaks in both RR and DET that were similar to
that of the all-events analysis. As mentioned above, this time period was almost
consistently dominated by events with a score of -10, indicating the most
violent, dramatic events; during this time, Twitter users may come together to
express grievances, bring about public awareness, ask for the help of other
countries, or just document the events. However, further semantic analysis is
needed to investigate these hypotheses.

The above-chance values for negative-event coupling are followed by a consistent
period of below-chance RR and a drastic drop in DET, co-occurring with the
increase in positive-event coupling. This again may reflect the shift in
attention to some positive events (e.g., aid, ceasefire agreements) or even a
decrease in hostilities. However, the present analyses should be complemented
with future analyses of semantic content to investigate these possible
explanations.

### Syria as Source and Target
For count of all events and social cohesion, the WCRQA plot for RR looked
largely the same as that of the target-only events, but DET showed a new early
peak in above-chance values. This suggests that the overall number of events
were more consistent from day to day when considering source events, as might be
expected given that individuals can only participate in a limited number of
events each day. Days that previously had a lower than average number of target
events may now have had a more near-average number of source and target events
because the reduced participation in target events allows for increased
participation and initiation of source events.

For positive events, we see a similar trend in RR as was observed in the
target-only dataset, with a more pronounced late peak. However, the previously
significant peaks in DET were no longer present. The addition of source events
may have thus increased the amount of recurrence but overshadowed the
trajectories of shared values found in the target-only events.

For negative events, the early peak in RR for the target-only filtering was
dramatically reduced compared to the target-only dataset, now spanning only two
windows. Similar to that of positive events, the late peak in DET was no longer
present. We hypothesize again that the addition of source events may have
obscured the true coupled relationship between the negative events and social
cohesion.

@@

After identifying the dataset of interest, researchers should think carefully
about how they will winnow down their data for analysis. The differences in our
findings when considering target-only or source-and-target filtering highlight
the importance of considering directionality when using event data. The parties
involved and the specific actions they take can dramatically influence results,
and it is critical to think carefully about which will be chosen and why.

This methodology should be of great interest to interdisciplinary scientists
investigating social mobilization. It has potential applications for gender
studies and the #MeToo movement and for the relationship between social protest
and repression during the Black Lives Matter protests. This methodological
approach can help us better understand the fundamental human social mechanisms
underlying social media and its influence on human behavior in real life. Given
that many governments enact repressive measures when they determine that
collective action by citizens is imminent, this methodology should help to
improve the predictive power of using social media data in estimating repressive
behavior. If social media messages are increasing in cohesion, this may
represent a threat to civil and political liberties of denizens engaging in
collective action, a unified reaction to sparse positive events, or an attempt
to decry the government’s attempt to distract from what may be going on.

## Conclusion
Considering the pervasiveness of social media, it is unsurprising that social
activists and protesters have utilized these platforms to support real-world
mobilization. Here, we quantitatively explored the connection between online
social cohesion and real-world events. In addition to introducing political
scientists to a family of powerful nonlinear dynamics analyses, our results have
serious implications for monitoring global movements. While state-run media in
regions of authoritarian rule are often biased and serve as propaganda for the
leader and central government, the self-organized nature of Twitter provides a
new avenue to gain insight into truths about uprisings and social mobilization.
By conceptualizing Twitter and real-world events as two elements of an
inextricably linked dynamical system, the levels of social cohesion within
social media platforms could be monitored for fluctuations indicating shifts in
power and peace. Social media mobilization may thus help alert the world to
citizens' actions and reactions that are otherwise suppressed by government
regimes, providing insight into events that may not otherwise be broadcast and
allowing for faster intervention.
